Random_Effects

### 2023-07-05 Code/BatchEffect.Rmd
Clarified models

```{r}
library(lme4)
two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
year_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
batch_effect <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
```

```{}
The code chunk above creates 4 models: no random effects, a year effect only, 
batch effect only, and both year and batch effect.
```

```{r}
anova(batch, two_effect)
```
```{}
This anova test determined that the two_effect model is an improvement over the
batch effect model. The test statistic is significant at alpha = 0.05 but not at
alpha = 0.01.
```

```{r}
anova(one_effect, two_effect)
```
```{}
In this anova test, it is clear that the two random effect model is superior to
having only the year effect.
```

```{r}
old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))

```

```{}
This section fits a two random effect model for each dataset.
```

>>>>

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
```
```{}
This loads in the datasets
```

```{r}
library(lme4)
men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))

men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))

men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))

men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))

```
```{}
This creates the year effect and the no year effect models
```

```{r}
men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, family = Gamma(link = "log"))
men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, family = Gamma(link = "log"))
men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
```
```{}
This creates the batch effect only models
```

```{r}
men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
```
```{}
This creates the year and batch effect models
```

```{r}
anova(men_pooled_batch, men_pooled_two_effect)
```
```{}
This anova test determined that the two_effect model is an improvement over the
batch effect model. The test statistic is significant at alpha = 0.05 but not at
alpha = 0.01.
```

```{r}
anova(men_pooled_year_gamma_log, men_pooled_two_effect)
```
```{}
In this anova test, it is clear that the two random effect model is superior to
having only the year effect.
```
